---
title: "Adipokine EWAS - Script 2: Leptin EWAS the in Leiden Longevity Study (LLS)"
author: "Lucy Sinke"
output: html_document
---

```{r echo=FALSE}
rm(list=ls())
```

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/exports/molepi/users/ljsinke/LLS/Lep_Scripts")
```

***

# Setup

***

Load packages

```{r message=FALSE}
library(SummarizedExperiment)
library(haven)
library(tidyverse)
library(ggpubr)
library(DNAmArray)
library(minfi)
library(limma)
```

***

# DNA Methylation Data

***

Whole blood DNA methylation was profiled by the Infinium HumanMethylation450 BeadChip. Preprocessing and quality control was performed in R.

DNA methylation is expressed as β values, which will be the response variable in all models. A `SummarizedExperiment` object containing this data, `betas`, is loaded in.

```{r}
load("../Shared_Data/LLS_betas.Rdata")
betas
```

This contains DNA methylation data for 732 individuals at 483,871 CpG sites.

***

# Phenotypes

***

### Leptin

Load in leptin (ng/mL) data from the LLS. 

* `llnr_id` - ID variable
* `age` - Age, a covariate
* `leptin` - Leptin levels

```{r}
file_name <- "LLS_IOP1-450kDNAmethylation-AdiponectinLeptinIl6_20201009.sav"

lls_cytokines <- read_sav(paste0("../Shared_Data/", file_name))

lls_cytokines <- lls_cytokines %>% 
  dplyr::select(llnr_id = LLnr, 
                age = Age, 
                leptin = Leptin) %>% 
  filter(!is.na(leptin))

summary(lls_cytokines)
```

***

### Limit of Detection

Remove values below LOD

```{r}
lls_cytokines <- lls_cytokines %>% 
  mutate(
    leptin = ifelse(lls_cytokines$leptin == 0, NA, leptin)
  )

summary(lls_cytokines$leptin)
summary(lls_cytokines)
```

***

### Normality

Check the Normality of the Leptin level distribution.

```{r warning=FALSE, message=FALSE}
ggdensity(lls_cytokines$leptin, 
          main = "Density plot of Leptin Levels",
          xlab = "Leptin")

ggqqplot(lls_cytokines$leptin)

shapiro.test(lls_cytokines$leptin)

summary(lls_cytokines$leptin)
```

***

### Natural Log Transformation

The distribution is negatively skewed, so it is natural log-transformed to improve Normality.

```{r warning=FALSE}
lls_cytokines$log_lep <- log(lls_cytokines$leptin)
```


***

### Outlier Removal

Any log(Leptin) values further than 3 IQR from the nearest quartile are excluded.

```{r}
summary(lls_cytokines$log_lep)

lower_limit <- quantile(lls_cytokines$log_lep, na.rm=TRUE)[[2]] - 
                      (3 * IQR(lls_cytokines$log_lep, na.rm=TRUE))
print(paste0("The lower limit is ", round(lower_limit,2)))

upper_limit <- quantile(lls_cytokines$log_lep, na.rm=TRUE)[[4]] +
                      (3 * IQR(lls_cytokines$log_lep, na.rm=TRUE))
print(paste0("The upper limit is ", round(upper_limit,2)))
```

One value outside of this range is excluded

```{r}
lls_cytokines <- lls_cytokines %>% 
  mutate(
     log_lep = ifelse(
       lls_cytokines$log_lep < lower_limit | lls_cytokines$log_lep > upper_limit, 
       NA, log_lep),
     leptin = ifelse(
       lls_cytokines$log_lep < lower_limit | lls_cytokines$log_lep > upper_limit, 
       NA, leptin)
)

summary(lls_cytokines$leptin)
summary(lls_cytokines)
```

***

### Final Normality Check

Check the Normality of the resulting distribution.

```{r warning=FALSE}
ggdensity(lls_cytokines$log_lep, 
          main = "Density plot of log-transformed Leptin",
          xlab = "log(Leptin)")

ggqqplot(lls_cytokines$log_lep)
shapiro.test(lls_cytokines$log_lep)
summary(lls_cytokines$log_lep)
```

Normality is visually improved

***

# IDs

***

ID values are stored for the samples with DNA methylation data.

```{r}
targets <- data.frame(uuid = rownames(colData(betas)),
                      run_id = colData(betas)$run_id)
```

***

# Covariates

***

### Age

Age is an a priori confounder and will be included in all models. It is stored with the cytokine data in `lls_cytokines`.

```{r}
summary(lls_cytokines$age)
```

***

### Sex

Sex is also an a priori confounder and is stored in the `colData` of the DNA methylation data. It is recoded as

* 0 for males
* 1 for females

```{r}
targets <- targets %>% 
  mutate(
    sex = factor(
      case_when(
         colData(betas)$sex == "male" ~ 0,
         colData(betas)$sex == "female" ~ 1
      ), labels = c("Male", "Female")))

xtabs(~targets$sex, addNA=TRUE)
```

***

### Smoking Status

There will be a smoking sensitivity analysis. 

Save smoking data, coded as:

* 0 for never smokers
* 1 for former smokers
* 2 for current smokers

```{r}
targets <- targets %>% 
  mutate(
     smoking = factor(
       case_when(
          colData(betas)$smoking == "non-smoker" ~ 0,
          colData(betas)$smoking == "former-smoker" ~ 1,
          colData(betas)$smoking == "current smoker" ~ 2
  ), labels = c("Never", "Former", "Current")))

xtabs(~targets$smoking, addNA=TRUE)
```

***

### Body Mass Index (BMI)

There will be a BMI sensitivity analysis.

BMI data is stored.

```{r}
targets$bmi <- colData(betas)$weight / ((colData(betas)$height/100)^2)

summary(targets$bmi)
```

***

### IDOL Cell counts

IDOL cell counts have been predicted from DNA methylation for LLS

```{r warning=FALSE, message=FALSE}
counts <- read_tsv("../Shared_Data/LLS_countsUCB.tsv")
colnames(counts) <- c("run_id", "CD8T", "CD4T", "NK", "Bcell", "Mono", "Gran")
head(counts)
```

These are merged with the rest of the covariate data

```{r}
targets <- left_join(targets, counts, by = "run_id")
```

***

### IDOL Extended Cell Counts

IDOL Extended cell counts have been predicted from DNA methylation data in the LLS

```{r}
load("../Shared_Data/LLS_deconvolution.RData")

LLS.prop <- LLS.prop %>% rownames_to_column(var="uuid")
LLS.prop <- LLS.prop %>% dplyr::select(
  uuid,
  Mono_ext = Mono, 
  NK_ext = NK, 
  Bas_ext = Bas, 
  Bmem_ext = Bmem, 
  Bnv_ext = Bnv, 
  CD4mem_ext = CD4mem, 
  CD4nv_ext = CD4nv,
  CD8mem_ext = CD8mem, 
  CD8nv_ext = CD8nv, 
  Eos_ext = Eos, 
  Neu_ext = Neu,
  Treg_ext = Treg
)

head(LLS.prop)
```

These are merged with the rest of the covariate data.

```{r}
targets <- left_join(targets, LLS.prop, by="uuid")
```

***

### Technical covariates

Technical effects are associated with the row number of the EPIC array. Additionally, the batch effect of plate number will be adjusted for. 

```{r}
targets$plate <- colData(betas)$sample_plate
targets$row <- colData(betas)$sentrix_position
```

Plate will be a fixed effect, but row number will take continuous values depending on the row on the 450k chip.

```{r}
xtabs(~targets$row)
targets <- targets %>% 
  mutate(
    row = as.numeric(substr(row, 3, 3)))
xtabs(~targets$row)
```

***

# Merging

***

The final structure of the covariate data:

```{r}
str(targets)
```

Load in the ID key for LLS

```{r echo=FALSE}
load("../Shared_Data/LLS_key.Rdata")
```

Keep only the IDs required

```{r}
lls_key <- lls_key %>% 
  dplyr::select(
    llnr_id, 
    uuid)
```

Add `uuid` to the cytokine data

```{r}
lls_cytokines <- left_join(
  lls_cytokines, 
  lls_key, 
  by="llnr_id")

str(lls_cytokines)
```

Since `targets` is derived from `methData`, it contains only observations where we also have methylation data. Therefore, we merge the other datasets into targets, excluding observations where beta values are missing.

```{r}
targets <- left_join(
  targets, 
  lls_cytokines, 
  by = "uuid")
```

Arrange columns in an intuitive order and remove `iop2_id`.

```{r}
targets <- targets %>% 
  dplyr::select(
    uuid, 
    log_lep, leptin,
    age, sex, 
    plate, row,
    CD8T, CD4T, NK, Bcell, Mono, Gran,
    smoking, bmi,
    Bas_ext, Bmem_ext, Bnv_ext, 
    CD4mem_ext, CD4nv_ext, CD8mem_ext, CD8nv_ext, 
    Eos_ext, Mono_ext, 
    Neu_ext, NK_ext, Treg_ext
    )
targets <- zap_labels(targets)
str(targets)
```

***

# Exclusion Criteria

***

### Sex Chromosomes

Probes from the X chromosome and the Y chromosome were excluded from the analysis. This excludes 11,263 probes from the analysis. 

```{r}
dim(betas)
betas <- betas[!(seqnames(betas) %in% c("chrX", "chrY")),]
dim(betas)
```

***

### Encode Blacklist Regions

Load in the CpGs within ENCODE blacklist regions

```{r}
load(
    "/exports/molepi/users/ljsinke/LLS/Shared_Data/ENCODE_Blacklist/ENCODEBlacklist_CpGomit-450K.RData")
```

These 5,889 probes are excluded from the methylation data

```{r}
betas
betas <- betas[!rownames(betas) %in% cpg_blacklist,]
betas
```

***

### Zhou probes

Load in Zhou probes recommendations using `DNAmArray`

```{r}
maskProbes <- read_tsv("../Shared_Data/Manifests/HM450.hg19.manifest.tsv.gz")
dim(maskProbes)
```

For general use, there are 60,466 probes to be excluded

```{r}
maskProbes <- maskProbes[maskProbes$MASK_general == TRUE,]$probeID
length(maskProbes)
```

Remove 54,887 of these from the LLS data

```{r}
betas
betas <- betas[!rownames(betas) %in% maskProbes,]
betas
```

***

### Outlier Removal

Any data points outside the range of `LQ - 3 x IQR` to `UQ + 3 x IQR` were excluded. This included 1,325,525 values that were converted to NA. 

```{r}
xtabs(~is.na(assay(betas)))
iqr_dnam <- apply(assay(betas), 1, function(x){
  iqr <- IQR(x, na.rm = TRUE)
  q1 <- quantile(x, na.rm=TRUE)[[2]]
  q3 <- quantile(x, na.rm=TRUE)[[4]]
  x <- ifelse((x <= q1 - (3*iqr) | x >= q3 + (3*iqr)), NA, x)
})
xtabs(~is.na(iqr_dnam))
dim(iqr_dnam)
```

We replace the `assay` of `methData` with the new beta values.

```{r}
assay(betas) <- t(iqr_dnam)
```

***

### Missingness

Keep only the CpGs where less than 5% of sample values are missing

This is all CpGs, representing high quality data

```{r}
perc_na <- rowSums(is.na(iqr_dnam))*100/ncol(iqr_dnam)
summary(perc_na)

betas <- betas[,perc_na <= 95]
betas
```

Keep only the samples where less than 5% of CpGs are missing

This is all samples, representing high quality data

```{r}
perc_na <- colSums(is.na(iqr_dnam))*100/nrow(iqr_dnam)
summary(perc_na)

betas <- betas[perc_na <= 95,]
betas
```

***

# Merging

***

Filter `targets` to only include samples in `betas`

```{r}
targets <- targets %>% 
  filter(uuid %in% rownames(colData(betas)))
```

We double check that samples are in the correct order in `targets` before replacing `colData`

```{r}
rownames(targets) <- targets$uuid
identical(rownames(targets), rownames(colData(betas)))
```

Add sample information to the `colData` of `methData` 

```{r}
colData(betas) <- DataFrame(targets)
```

***

# Complete Data

***

### Base Model

The EWAS can only be performed on individuals with observed values for all covariates

```{r}
dim(betas)
betas <- subset(betas, 
                select = rowSums(is.na(colData(betas)[, c(1:13)])) == 0)
dim(betas)
```

***

### Smoking Sensitivity Analysis

Exclude those with missing data for smoking status.

```{r}
betas_smoke <- subset(betas, 
                      select = rowSums(is.na(colData(betas)[, c(1:14)])) == 0)
dim(betas_smoke)
```

***

### BMI Sensitivity Analysis

Save data with complete BMI information.

```{r}
betas_bmi <- subset(betas, 
                    select = rowSums(is.na(colData(betas)[, c(1:13,15)])) == 0)
dim(betas_bmi)
```

***

### IDOl Extended Sensitivity Analysis

Save data with extended cell count data.

```{r}
betas_idol <- subset(betas, 
                     select = rowSums(is.na(colData(betas)[, c(1:13,16)])) == 0)
dim(betas_idol)
```

***

# Save Covariates

***

Save targets

```{r}
save(targets, file = "../Lep_Data/Cohort_Results/LLS/targets_lep.Rdata")
```

***

# EWAS

***

### Base Model

The base model for the Leptin EWAS is:

DNAm ~ log_lep + age + sex + CD8T + CD4T + NK + Bcell + Mono + plate + row

* `age`, `row`, `CD4T`, `CD8T`, `NK`, `Bcell`, and `Mono` are all added as continuous fixed effects
* `sex` and `plate` added as a categorical fixed effect

Save the formula for the model

```{r}
metadata(betas)$formula <- ~log_lep + age + sex + CD8T + CD4T + NK + Bcell + Mono + plate + row
```

Create a design matrix

```{r}
design <- model.matrix(metadata(betas)$formula, 
                       data=colData(betas))
head(design)
```

Fit the model

```{r}
fit <- lmFit(assay(betas), design)
```

Store the effect sizes, SE, t-statistics, p-values, and sample size for `log_lep`

```{r}
beta <- fit$coefficients[, 2]
SE <- fit$stdev.unscaled[, 2] * fit$sigma
t <- beta / SE
p <- 2 * pt(-abs(t), fit$df.residual)
N <- ncol(design) + fit$df.residual
```

Adjust for multiple testing using FDR

```{r}
padj <- p.adjust(p, method="fdr")
```

Save the mean and standard deviation for each probe

```{r}
cpg_mean <- apply(assay(betas), 1, mean)
cpg_se <- apply(assay(betas), 1, sd)
```

Save these results in a data frame, `limma_base` 

```{r}
limma_base <- data.frame(
  cpg = rownames(fit$coefficients), 
  cpg_mean, 
  cpg_se,
  beta, 
  SE, 
  p, 
  padj, 
  N)
```

Save the results

```{r}
save(limma_base, file="../Lep_Data/Cohort_Results/LLS/LLS_ln-lep_base.Rdata")
```

***

### Smoking Sensitivity Analysis

The first extended model for the Leptin EWAS is:

DNAm ~ log_lep + age + sex + **smoking** + CD8T + CD4T + NK + Bcell + Mono + plate + row

* `age`, `row`, `CD4T`, `CD8T`, `NK`, `Bcell`, and `Mono` are all added as continuous fixed effects
* `sex`, `plate`, and `smoking` are added as categorical fixed effects

Save the formula for the model

```{r}
metadata(betas_smoke)$formula <- ~log_lep + age + sex + smoking + CD8T + CD4T + NK + Bcell + Mono + plate + row
```

Create a design matrix

```{r}
design <- model.matrix(metadata(betas_smoke)$formula, 
                       data=colData(betas_smoke))
head(design)
```

Fit model

```{r}
fit <- lmFit(assay(betas_smoke), design)
```

Store the effect sizes, SE, t-statistics, p-values, and sample size for `log_lep`.

```{r}
beta <- fit$coefficients[, 2]
SE <- fit$stdev.unscaled[, 2] * fit$sigma
t <- beta / SE
p <- 2 * pt(-abs(t), fit$df.residual)
N <- ncol(design) + fit$df.residual
```

Adjust for multiple testing using FDR

```{r}
padj <- p.adjust(p, method="fdr")
```

Save these results in a data frame, `limma_base`

```{r}
limma_base <- data.frame(
  cpg = rownames(fit$coefficients), 
  beta, 
  SE, 
  p, 
  padj, 
  N)
```

Save the results.

```{r}
save(limma_base, file="../Lep_Data/Cohort_Results/LLS/LLS_ln-lep_ext1.Rdata")
```

***

### BMI Sensitivity Analysis

The second extended model for the Leptin EWAS is:

DNAm ~ log_lep + age + sex + **BMI** + CD8T + CD4T + NK + Bcell + Mono + plate + row

* `age`, `row`, `bmi`, `CD4T`, `CD8T`, `NK`, `Bcell`, and `Mono` are all added as continuous fixed effects
* `sex` and `plate` are added as categorical fixed effects

Save the formula for the model

```{r}
metadata(betas_bmi)$formula <- ~log_lep + age + sex + bmi + CD8T + CD4T + NK + Bcell + Mono + plate + row
```

Create a design matrix

```{r}
design <- model.matrix(metadata(betas_bmi)$formula, 
                       data=colData(betas_bmi))
head(design)
```

Fit model

```{r}
fit <- lmFit(assay(betas_bmi), design)
```

Store the effect sizes, SE, t-statistics, p-values, and sample size for `log_lep`.

```{r}
beta <- fit$coefficients[, 2]
SE <- fit$stdev.unscaled[, 2] * fit$sigma
t <- beta / SE
p <- 2 * pt(-abs(t), fit$df.residual)
N <- ncol(design) + fit$df.residual
```

Adjust for multiple testing using FDR

```{r}
padj <- p.adjust(p, method="fdr")
```

Save these results in a data frame, `limma_base`

```{r}
limma_base <- data.frame(
  cpg = rownames(fit$coefficients), 
  beta, 
  SE, 
  p, 
  padj, 
  N)
```

Save the results.

```{r}
save(limma_base, file="../Lep_Data/Cohort_Results/LLS/LLS_ln-lep_ext2.Rdata")
```

***

### IDOL Extended Sensitivity Analysis

The third extended model for the Leptin EWAS is:

DNAm ~ log_lep + age + sex + Bas + Bmem + Bnv + CD4mem + CD4nv + CD8mem + CD8nv + Eos + Mono + NK + Treg + plate + row

* `age`, `row`, `Bas`, `Bmem`, `Bnv`, `CD4mem`, `CD4nv`, `CD8mem`, `CD8nv`, `Eos`, `Mono`, `NK`, and `Treg` are all added as continuous fixed effects
* `sex` and `plate` are added as categorical fixed effects

We save the formula for the model.

```{r}
metadata(betas_idol)$formula <- ~log_lep + age + sex + plate + row + 
  Bas_ext + Bmem_ext + Bnv_ext + 
  CD4mem_ext + CD4nv_ext + CD8mem_ext + CD8nv_ext + 
  Eos_ext + Mono_ext + NK_ext + Treg_ext 
```

Create a design matrix

```{r}
design <- model.matrix(metadata(betas_idol)$formula, 
                       data=colData(betas_idol))
head(design)
```

Fit model

```{r}
fit <- lmFit(assay(betas_idol), design)
```

Store the effect sizes, SE, t-statistics, p-values, and sample size for `log_lep`.

```{r}
beta <- fit$coefficients[, 2]
SE <- fit$stdev.unscaled[, 2] * fit$sigma
t <- beta / SE
p <- 2 * pt(-abs(t), fit$df.residual)
N <- ncol(design) + fit$df.residual
```

Adjust for multiple testing using FDR

```{r}
padj <- p.adjust(p, method="fdr")
```

Save these results in a data frame, `limma_base`

```{r}
limma_base <- data.frame(
  cpg = rownames(fit$coefficients), 
  beta, 
  SE, 
  p, 
  padj, 
  N)
```

Save the results.

```{r}
save(limma_base, file="../Lep_Data/Cohort_Results/LLS/LLS_ln-lep_ext3.Rdata")
```































