---
title: "Adipokine EWAS - Script 3: Saving Cohort Results"
author: "Lucy Sinke"
output: html_document
---

```{r echo=FALSE}
rm(list=ls())
```

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/exports/molepi/users/ljsinke/LLS/Adi_Scripts')
```

***

# Setup

Load packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

***

# Adiponectin EWAS - Base Model

Our base model, as specified in the common analysis plan, adjusts for `age`, `sex`, `predicted cell counts`, and `technical covariates`.

***

### Leiden Longevity Study (LLS)

Read in the LLS results

```{r}
load('../Adi_Data/Cohort_Results/LLS/LLS_ln-adi_base.Rdata')
limma_base <- limma_base %>% 
  dplyr::select(
    cpg, 
    beta_lls = beta, 
    SE_lls = SE, 
    p_lls = p, 
    N_lls = N)
```

***

### KORA

```{r}
df <- read_csv2('../Adi_Data/Cohort_Results/KORA/KORA_ln-adiponectin_base.csv')

df <- df %>% 
  dplyr::select(
    cpg = cg, 
    beta_kora = beta,
    SE_kora = se, 
    p_kora = p, 
    N_kora = `N obs`)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

### TwinsUK (TUK)

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/TUK/TUK_ln-adiponectin_base.csv')

df <- df %>% dplyr::select(
  cpg = CpG, 
  beta_tuk = beta,
  SE_tuk = SE, 
  p_tuk = p, 
  N_tuk = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

# LifeLines Deep (LLD)

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/LLD/LLD_ln-adiponectin_base.csv')

df <- df %>% 
  dplyr::select(
    cpg = CpG, 
    beta_lld = BETA,
    SE_lld = SE, 
    p_lld = P_VAL, 
    N_lld = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

# SHIP

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/SHIP/SHIP_ln-adiponectin_base.csv.gz')

df <- df %>% 
  dplyr::select(
    cpg = CpG, 
    beta_ship = beta,
    SE_ship = SE, 
    p_ship = P, 
    N_ship = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***


# GOLDN

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/GOLDN/GOLDN_ln-adiponectin_base.csv')

df <- df %>% 
  dplyr::select(
    cpg, 
    beta_goldn = Estimate,
    SE_goldn = Stderr, 
    p_goldn = ProbZ) %>% 
  mutate(N_goldn = 994)

limma_base <- full_join(
  limma_base, df, 
  by = "cpg")
```

***

### Save ALL

```{r}
limma_base <- limma_base %>% 
  dplyr::select(
    cpg, 
    starts_with('beta'), 
    starts_with('SE'),
    starts_with('p'), 
    starts_with('N'))

save(limma_base, file='../Adi_Data/Cohort_Results/ALL_ln-adi_base.Rdata')
```

***

# Adiponectin EWAS - Smoking Sensitivity Analysis

***

### Leiden Longevity Study (LLS)

Read in the LLS results

```{r}
load('../Adi_Data/Cohort_Results/LLS/LLS_ln-adi_ext1.Rdata')
limma_base <- limma_base %>% 
  dplyr::select(
    cpg, 
    beta_lls = beta, 
    SE_lls = SE, 
    p_lls = p, 
    N_lls = N)
```

***

### KORA

```{r}
df <- read_csv2('../Adi_Data/Cohort_Results/KORA/KORA_ln-adiponectin_ext1.csv')

df <- df %>% 
  dplyr::select(
    cpg = cg, 
    beta_kora = beta,
    SE_kora = se, 
    p_kora = p, 
    N_kora = `N obs`)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

### TwinsUK (TUK)

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/TUK/TUK_ln-adiponectin_ext1.csv')

df <- df %>% dplyr::select(
  cpg = CpG, 
  beta_tuk = beta,
  SE_tuk = SE, 
  p_tuk = p, 
  N_tuk = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

# LifeLines Deep (LLD)

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/LLD/LLD_ln-adiponectin_ext1.csv')

df <- df %>% 
  dplyr::select(
    cpg = CpG, 
    beta_lld = BETA,
    SE_lld = SE, 
    p_lld = P_VAL, 
    N_lld = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

# SHIP

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/SHIP/SHIP_ln-adiponectin_ext1.csv.gz')

df <- df %>% 
  dplyr::select(
    cpg = CpG, 
    beta_ship = beta,
    SE_ship = SE, 
    p_ship = P, 
    N_ship = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

# GOLDN

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/GOLDN/GOLDN_ln-adiponectin_ext1.csv')

df <- df %>% 
  dplyr::select(
    cpg, 
    beta_goldn = Estimate,
    SE_goldn = Stderr, 
    p_goldn = ProbZ) %>% 
  mutate(N_goldn = 994)

limma_base <- full_join(
  limma_base, df, 
  by = "cpg")
```

***

### Save ALL

```{r}
limma_base <- limma_base %>% 
  dplyr::select(
    cpg, 
    starts_with('beta'), 
    starts_with('SE'),
    starts_with('p'), 
    starts_with('N'))

save(limma_base, file='../Adi_Data/Cohort_Results/ALL_ln-adi_ext1.Rdata')
```

***

# Adiponectin EWAS - BMI Sensitivity Analysis

***

### Leiden Longevity Study (LLS)

Read in the LLS results

```{r}
load('../Adi_Data/Cohort_Results/LLS/LLS_ln-adi_ext2.Rdata')
limma_base <- limma_base %>% 
  dplyr::select(
    cpg, 
    beta_lls = beta, 
    SE_lls = SE, 
    p_lls = p, 
    N_lls = N)
```

***

### KORA

```{r}
df <- read_csv2('../Adi_Data/Cohort_Results/KORA/KORA_ln-adiponectin_ext2.csv')

df <- df %>% 
  dplyr::select(
    cpg = cg, 
    beta_kora = beta,
    SE_kora = se, 
    p_kora = p, 
    N_kora = `N obs`)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

### TwinsUK (TUK)

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/TUK/TUK_ln-adiponectin_ext2.csv')

df <- df %>% dplyr::select(
  cpg = CpG, 
  beta_tuk = beta,
  SE_tuk = SE, 
  p_tuk = p, 
  N_tuk = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

# LifeLines Deep (LLD)

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/LLD/LLD_ln-adiponectin_ext2.csv')

df <- df %>% 
  dplyr::select(
    cpg = CpG, 
    beta_lld = BETA,
    SE_lld = SE, 
    p_lld = P_VAL, 
    N_lld = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

# SHIP

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/SHIP/SHIP_ln-adiponectin_ext2.csv.gz')

df <- df %>% 
  dplyr::select(
    cpg = CpG, 
    beta_ship = beta,
    SE_ship = SE, 
    p_ship = P, 
    N_ship = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***
# GOLDN

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/GOLDN/GOLDN_ln-adiponectin_ext2.csv')

df <- df %>% 
  dplyr::select(
    cpg, 
    beta_goldn = Estimate,
    SE_goldn = Stderr, 
    p_goldn = ProbZ) %>% 
  mutate(N_goldn = 994)

limma_base <- full_join(
  limma_base, df, 
  by = "cpg")
```

***

### Save ALL

```{r}
limma_base <- limma_base %>% 
  dplyr::select(
    cpg, 
    starts_with('beta'), 
    starts_with('SE'),
    starts_with('p'), 
    starts_with('N'))

save(limma_base, file='../Adi_Data/Cohort_Results/ALL_ln-adi_ext2.Rdata')
```

***


***

# Adiponectin EWAS - epidish sensitivity analysis

***

### Leiden Longevity Study (LLS)

Read in the LLS results

```{r}
load('../Adi_Data/Cohort_Results/LLS/LLS_ln-adi_ext3.Rdata')
limma_base <- limma_base %>% 
  dplyr::select(
    cpg, 
    beta_lls = beta, 
    SE_lls = SE, 
    p_lls = p, 
    N_lls = N)
```

***

### KORA

```{r}
df <- read_csv2('../Adi_Data/Cohort_Results/KORA/KORA_ln-adiponectin_ext3.csv')

df <- df %>% 
  dplyr::select(
    cpg = cg, 
    beta_kora = beta,
    SE_kora = se, 
    p_kora = p, 
    N_kora = N_obs)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

# LifeLines Deep (LLD)

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/LLD/LLD_ln-adiponectin_ext3.csv')

df <- df %>% 
  dplyr::select(
    cpg = CpG, 
    beta_lld = BETA,
    SE_lld = SE, 
    p_lld = P_VAL, 
    N_lld = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

# SHIP

```{r}
df <- read_csv('../Adi_Data/Cohort_Results/SHIP/SHIP_ln-adiponectin_ext3.gz')

df <- df %>% 
  dplyr::select(
    cpg = CpG, 
    beta_ship = beta,
    SE_ship = SE, 
    p_ship = P, 
    N_ship = N)

limma_base <- full_join(
  limma_base, df, 
  by = 'cpg')
```

***

### Save ALL

```{r}
limma_base <- limma_base %>% 
  dplyr::select(
    cpg, 
    starts_with('beta'), 
    starts_with('SE'),
    starts_with('p'), 
    starts_with('N'))

save(limma_base, file='../Adi_Data/Cohort_Results/ALL_ln-adi_ext3.Rdata')
```

***













