---
title: "Adiponectin EWAS - Script 6: EWAS QC Post-bacon"
author: "Lucy Sinke"
output: html_document
---

```{r echo=FALSE}
rm(list=ls())
```


```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/exports/molepi/users/ljsinke/LLS/Adi_Scripts")
```

***

# Setup

Load packages

```{r message=FALSE}
library(ggpubr)
library(DNAmArray)
library(lattice)
library(ggrepel)
library(tidyverse)
```

***

### Load probe annotation

For 450K data:

```{r warning=FALSE, message = FALSE}
anno <- read_tsv("../Shared_Data/Manifests/HM450.hg19.manifest.tsv.gz")

anno <- anno %>% 
  dplyr::select(
    cpg = probeID,
    chr = CpG_chrm,
    start = CpG_beg,
    MASK_general)
```

For EPIC data:

```{r}
anno_EPIC <- read_tsv("../Shared_Data/Manifests/EPIC.hg19.manifest.tsv.gz")

anno_EPIC <- anno_EPIC %>% 
  dplyr::select(
    cpg = probeID,
    chr = CpG_chrm,
    start = CpG_beg
  )
```

***

# Adiponectin EWAS - Base Model

***

### Load data

This contains data on:

* `cpg` - probe ID
* `beta_` - effect size reported by each cohort
* `SE_` - standard error reported by each cohort
* `p_`- unadjusted p-value reported by each cohort
* `N_` - sample size of each cohort

```{r}
load('../Adi_Data/Processing/ALL-3-BACON_ln-adi_base.Rdata')
bacon_df[1:5, 1:5]
```

***

### Save Results

Create function

```{r}
save_results <- function(cohort){
  assign(paste0("df_", cohort),
         bacon_df %>% 
           dplyr::select(
             cpg,
             beta = paste0("beta_", cohort),
             se = paste0("SE_", cohort),
             p = paste0("p_", cohort),
             N = paste0("N_", cohort)
           ) %>% 
           mutate(
             beta = as.numeric(beta),
             se = as.numeric(se), 
             p = as.numeric(p),
             N = as.numeric(N) 
           ) %>% 
           filter(N>50),
         envir=.GlobalEnv)
}
```

***

LLS

```{r}
save_results("lls")
write_tsv(df_lls, file="../Adi_Data/Processing/Meta_Input/OUT_adi_base-LLS.tsv")
```

***

KORA

```{r}
save_results("kora")
write_tsv(df_kora, file="../Adi_Data/Processing/Meta_Input/OUT_adi_base-KORA.tsv")
```

***

TwinsUK

```{r}
save_results("tuk")
write_tsv(df_tuk, file="../Adi_Data/Processing/Meta_Input/OUT_adi_base-TUK.tsv")
```

***

LifeLines Deep

```{r}
save_results("lld")
write_tsv(df_lld, file="../Adi_Data/Processing/Meta_Input/OUT_adi_base-LLD.tsv")
```

***

SHIP

```{r}
save_results("ship")
write_tsv(df_ship, file="../Adi_Data/Processing/Meta_Input/OUT_adi_base-SHIP.tsv")
```

***

GOLDN

```{r}
save_results("goldn")
write_tsv(df_goldn, file="../Adi_Data/Processing/Meta_Input/OUT_adi_base-GOLDN.tsv")
```


### Top Hits

***

```{r}
df_lls <- left_join(df_lls, anno, by = "cpg")
head(df_lls %>% arrange(p))
```

***

KORA

```{r}
df_kora <- left_join(df_kora, anno, by = "cpg")
head(df_kora %>% arrange(p))
```

***

TwinsUK

```{r}
df_tuk <- left_join(df_tuk, anno, by = "cpg")
head(df_tuk %>% arrange(p))
```

***

LifeLines Deep

```{r}
df_lld <- left_join(df_lld, anno, by = "cpg")
head(df_lld %>% arrange(p))
```

***

SHIP

```{r}
df_ship <- left_join(df_ship, anno, by = "cpg")
head(df_ship %>% arrange(p))
```

***

GOLDN

```{r}
df_goldn <- left_join(df_goldn, anno, by="cpg")
head(df_goldn %>% arrange(p))
```


***


## Quality Control Plots

For all cohorts, we want to plot:

* QQ plots side by side
* Volcano plots side by side
* Manhattan plots side by side
* Boxplots of effect sizes
* Boxplots of SEs

***

### QQ plots 

Create function

```{r}
draw_qq <- function(cohort){
  n <- nrow(get(paste0("df_",cohort)))
  exp_p <- -log10((rank(get(paste0("df_", cohort))$p,
                        ties.method="first") - 0.5) / n)
  obs_p <- -log10(get(paste0("df_", cohort))$p)
  df <- data.frame(
    e=exp_p,
    o=obs_p
  )
  
  assign(paste0("plot_", cohort), 
         df %>% 
            ggplot() +
            geom_abline(
              aes(
                intercept=0, 
                slope=1), 
              color="#D62839") +
           geom_point(
             aes(
               x=e, 
               y=o), 
             alpha=0.5, 
             size=1, 
             color="#1B2021") +
           ylim(0,NA) + 
           xlim(0,NA) +
           ggtitle(toupper(cohort)) +
           xlab("Expected") +
           ylab("Observed") +
  theme(
    axis.text = element_text(
       size=9, 
       color="#1B2021"),
    axis.title = element_text(
      size=11, 
      hjust=0.5, 
      color="#1B2021"),
    plot.title = element_text(
      size=16, 
      hjust=0.5,
      face="bold", 
      color = "#548687"),
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021",
      fill=NA),
    panel.grid.major = element_line(
      color="grey95"),
    panel.grid.minor = element_line(
      color="grey95"),
    plot.background = element_rect(
      fill="white")),
  envir = .GlobalEnv)
}
```

***

LLS

```{r}
draw_qq("lls")
```

***

KORA

```{r}
draw_qq("kora")
```

***

TwinsUK

```{r}
draw_qq("tuk")
```

***

LifeLines Deep

```{r}
draw_qq("lld")
```

***

SHIP

```{r}
draw_qq("ship")
```

***

GOLDN

```{r}
draw_qq("goldn")
```

***

Draw

```{r out.width="100%"}
ggarrange(plot_lls, 
          plot_kora, 
          plot_tuk,
          plot_lld,
          plot_ship,
          plot_goldn,
          ncol = 3, nrow = 2)
```

***

### Volcano Plots

***

Make scale

```{r}
min <- as.numeric(-max(c(
  abs(bacon_df$beta_lls), 
  abs(bacon_df$beta_kora),
  abs(bacon_df$beta_tuk),
  abs(bacon_df$beta_lld),
  abs(bacon_df$beta_ship),
  abs(bacon_df$beta_goldn)),
  na.rm=TRUE) - 0.005)

max <- as.numeric(max(c(
  abs(bacon_df$beta_lls), 
  abs(bacon_df$beta_kora),
  abs(bacon_df$beta_tuk),
  abs(bacon_df$beta_lld),
  abs(bacon_df$beta_ship),
  abs(bacon_df$beta_goldn)),
  na.rm=TRUE) + 0.005)

p_max <- as.numeric(-log10(min(c(
  bacon_df$p_lls,
  bacon_df$p_kora,
  bacon_df$p_tuk,
  bacon_df$p_lld,
  bacon_df$p_ship,
  bacon_df$p_goldn),
  na.rm=TRUE))) + 2
```

Create function

```{r}
draw_volcano <- function(cohort){
  sig_df <- get(paste0("df_", cohort)) %>% 
    mutate(
      padj = p.adjust(p, method="fdr")
    ) %>% 
    filter(
      padj <= 0.05
    )
  
  if(nrow(sig_df) >= 1){
    sig_limit <- max(sig_df$p)
  } else {
    sig_limit <- 10E-07
  }
  
  assign(paste0("plot_", cohort), 
         get(paste0('df_', cohort)) %>% 
           ggplot(aes(
             x = beta,
             y = -log10(p)
           )) +
           geom_hline(
             yintercept = -log10(sig_limit),
             linetype = "dashed",
             size = 0.5,
             color = "#ACAFAF"
           ) +
           geom_vline(
             xintercept = -0.05,
             linetype = "dashed",
             size = 0.5,
             color = "#ACAFAF"
           ) +
           geom_vline(
             xintercept = 0.05,
             linetype = "dashed",
             size = 0.5,
             color = "#ACAFAF"
           ) +
         geom_point(
           color = ifelse(get(paste0('df_', cohort))$p <= sig_limit, 
                          "#D62839", "#ACAFAF"),
           size = 0.8, 
         ) +
           xlim(min, max) +
           ylim(0, p_max) +
           ggtitle(toupper(cohort)) +
           ylab(bquote(-log[10]~"p")) +
           xlab("Effect Size") +
  theme(
    axis.text = element_text(
       size=9, 
       color="#1B2021"),
    axis.title = element_text(
      size=11, 
      hjust=0.5, 
      color="#1B2021"),
    plot.title = element_text(
      size=16, 
      hjust=0.5,
      face="bold", 
      color = "#548687"),
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021",
      fill=NA),
    panel.grid.major = element_line(
      color="grey95"),
    panel.grid.minor = element_line(
      color="grey95"),
    plot.background = element_rect(
      fill="white")),
  envir = .GlobalEnv)
}
```

***

LLS

```{r}
draw_volcano("lls")
```

***

KORA

```{r}
draw_volcano("kora")
```

***

TwinsUK

```{r}
draw_volcano("tuk")
```

***

LifeLines Deep

```{r}
draw_volcano("lld")
```

***

SHIP

```{r}
draw_volcano("ship")
```

***

GOLDN

```{r}
draw_volcano("goldn")
```

***

Draw

```{r out.width="100%"}
ggarrange(
  plot_lls, 
  plot_kora, 
  plot_tuk,
  plot_lld,
  plot_ship,
  plot_goldn,
  ncol = 3, nrow = 2)
```

***

### Manhattan plots

***

Create function

```{r}
draw_man <- function(cohort){
  # Get sig limit
  sig_df <- get(paste0("df_", cohort)) %>% 
    mutate(
      padj = p.adjust(p, method="fdr")
    ) %>% 
    filter(
      padj <= 0.05
    )
  
  if(nrow(sig_df) >= 1){
    sig_limit <- max(sig_df$p)
  } else {
    sig_limit <- 10E-07
  }
  
  # Create data frame
  limma_man <- get(paste0("df_", cohort)) %>% 
    arrange(
      chr, start
    )
  
  # Initialize variables
  limma_man$start_cum <- NA
  s <- 0
  nbp <- c()
  
  # Loop
  for(i in unique(limma_man$chr)){
    nbp[i] <- max(
      limma_man[limma_man$chr == i,]$start
    )
    limma_man[limma_man$chr == i, "startcum"] <- limma_man[limma_man$chr == i, 
                                                           "start"] + s
    s <- s + nbp[i]
  }
  
  # Set axis
  axis.set <- limma_man %>% 
    group_by(chr) %>% 
    summarize(center = (max(startcum) + min(startcum)) / 2)
  
  # Draw plot
  assign(paste0('plot_', cohort),
         ggplot(
           limma_man,
           aes(
             x=startcum,
             y=-log10(p),
             color=as.factor(chr))) +
           geom_point(
             size=0.8
           ) + 
           geom_hline(
             yintercept = -log10(sig_limit),
             color = "#D62839",
             size = 0.5, 
             linetype = "dashed"
           ) +
           scale_x_continuous(
             label = axis.set$chr,
             breaks = axis.set$center
           ) +
           scale_y_continuous(
             expand = c(0,0),
             limits = c(0, p_max)
           ) +
           scale_size_continuous(
             range = c(0.5, 3)
           ) +
           scale_color_manual(
             values = rep(c("#ACAFAF", "#548687"), 11)
           ) +
           xlab("") +
           ylab(bquote(-log[10]~"p")) +
           ggtitle(toupper(cohort)) +
           theme( 
              legend.position = "none",
              panel.background = element_rect(fill="white"),
              panel.border = element_rect(color="#1B2021", fill=NA),
              panel.grid.major.x = element_blank(),
              panel.grid.minor.x = element_blank(),
              axis.text.y = element_text(size = 9, color = '#1B2021'),
              axis.text.x = element_text(angle = 90, size = 4, 
                               vjust = 0.5, color = '#1B2021'),
              axis.title = element_text(size=11, color = '#1B2021'),
              plot.title = element_text(size = 12, hjust=0.5,
                              color = '#548687', face = 'bold')),
         envir=.GlobalEnv)
  
  
}
```

***

LLS

```{r}
draw_man("lls")
```

***

KORA

```{r}
draw_man("kora")
```

***

TwinsUK

```{r}
draw_man("tuk")
```

***

LifeLines Deep

```{r}
draw_man("lld")
```

***

SHIP

```{r}
draw_man("ship")
```

***

GOLDN

```{r}
draw_man("goldn")
```

***

Draw

```{r out.width="100%"}
ggarrange(
  plot_lls, 
  plot_kora, 
  plot_tuk,
  plot_lld,
  plot_ship,
  plot_goldn,
  ncol = 3, nrow = 2)
```

***

### Boxplots

***

Make plot of effect size distributions in each cohort

```{r}
bacon_df <- bacon_df %>% 
  mutate(
    beta_LLS = as.numeric(beta_lls),
    beta_KORA = as.numeric(beta_kora),
    beta_TwinsUK = as.numeric(beta_tuk),
    beta_LLD = as.numeric(beta_lld),
    beta_SHIP = as.numeric(beta_ship),
    beta_GOLDN = as.numeric(beta_goldn)
  )

beta_df <- pivot_longer(
  data = bacon_df,
  cols = c("beta_LLS", "beta_KORA", "beta_TwinsUK", 
           "beta_LLD", "beta_SHIP", "beta_GOLDN"),
  names_to = "cohort",
  names_prefix = "beta_",
  values_to = "beta",
  values_drop_na = TRUE
)

beta_df %>% 
  ggplot(aes(x = cohort, y = beta, fill = cohort)) +
  scale_fill_manual(values = c('#e2455f', '#66cec8', '#eaa87d',
                               '#f9f5e0', '#794f7a', '#5FA052')) +
  geom_boxplot(alpha = 0.8) +
  ggtitle("Effect size distributions by cohort") +
  ylab("\u03b2") + xlab("Cohort") +
  theme( 
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021", fill=NA),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(size = 12, color = "#1B2021"),
    axis.text.x = element_text(size = 12, color = "#1B2021"),
    axis.title = element_text(size=11, color = "#1B2021"),
    plot.title = element_text(size = 16, hjust=0.5,
                              color = "#548687", face = "bold"))
```

***

Make plot of SE distribution in each cohort

```{r}
bacon_df <- bacon_df %>% 
  mutate(
    SE_LLS = as.numeric(SE_lls),
    SE_KORA = as.numeric(SE_kora),
    SE_TwinsUK = as.numeric(SE_tuk),
    SE_LLD = as.numeric(SE_lld),
    SE_SHIP = as.numeric(SE_ship),
    SE_GOLDN = as.numeric(SE_goldn)
  )

SE_df <- pivot_longer(
  data = bacon_df,
  cols = c("SE_LLS", "SE_KORA", "SE_TwinsUK", 
           "SE_LLD", "SE_SHIP", "SE_GOLDN"),
  names_to = "cohort",
  names_prefix = "SE_",
  values_to = "SE",
  values_drop_na = TRUE
)

SE_df %>% 
  ggplot(aes(x = cohort, y = SE, fill = cohort)) +
  scale_fill_manual(values = c('#e2455f', '#66cec8', '#eaa87d',
                               '#f9f5e0', '#794f7a', '#5FA052')) +
  geom_boxplot(alpha = 0.6) +
  ggtitle("SE distributions by cohort") +
  ylab("SE") + xlab("Cohort") +
  theme( 
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021", fill=NA),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(size = 12, color = "#1B2021"),
    axis.text.x = element_text(size = 12, color = "#1B2021"),
    axis.title = element_text(size=11, color = "#1B2021"),
    plot.title = element_text(size = 16, hjust=0.5,
                              color = "#548687", face = "bold"))
```


***

# Adiponectin EWAS - Smoking Sensitivity Analysis

***

### Load data

This contains data on:

* `cpg` - probe ID
* `beta_` - effect size reported by each cohort
* `SE_` - standard error reported by each cohort
* `p_`- unadjusted p-value reported by each cohort
* `N_` - sample size of each cohort

```{r}
load('../Adi_Data/Processing/ALL-3-BACON_ln-adi_ext1.Rdata')
bacon_df[1:5, 1:5]
```

***

### Save Results

Create function

```{r}
save_results <- function(cohort){
  assign(paste0("df_", cohort),
         bacon_df %>% 
           dplyr::select(
             cpg,
             beta = paste0("beta_", cohort),
             se = paste0("SE_", cohort),
             p = paste0("p_", cohort),
             N = paste0("N_", cohort)
           ) %>% 
           mutate(
             beta = as.numeric(beta),
             se = as.numeric(se), 
             p = as.numeric(p),
             N = as.numeric(N) 
           ) %>% 
           filter(N>50),
         envir=.GlobalEnv)
}
```

***

LLS

```{r}
save_results("lls")
write_tsv(df_lls, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext1-LLS.tsv")
```

***

KORA

```{r}
save_results("kora")
write_tsv(df_kora, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext1-KORA.tsv")
```

***

TwinsUK

```{r}
save_results("tuk")
write_tsv(df_tuk, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext1-TUK.tsv")
```

***

LifeLines Deep

```{r}
save_results("lld")
write_tsv(df_lld, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext1-LLD.tsv")
```

***

SHIP

```{r}
save_results("ship")
write_tsv(df_ship, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext1-SHIP.tsv")
```

***

GOLDN

```{r}
save_results("goldn")
write_tsv(df_goldn, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext1-GOLDN.tsv")
```


### Top Hits

***

```{r}
df_lls <- left_join(df_lls, anno, by = "cpg")
head(df_lls %>% arrange(p))
```

***

KORA

```{r}
df_kora <- left_join(df_kora, anno, by = "cpg")
head(df_kora %>% arrange(p))
```

***

TwinsUK

```{r}
df_tuk <- left_join(df_tuk, anno, by = "cpg")
head(df_tuk %>% arrange(p))
```

***

LifeLines Deep

```{r}
df_lld <- left_join(df_lld, anno, by = "cpg")
head(df_lld %>% arrange(p))
```

***

SHIP

```{r}
df_ship <- left_join(df_ship, anno, by = "cpg")
head(df_ship %>% arrange(p))
```

***

GOLDN

```{r}
df_goldn <- left_join(df_goldn, anno, by="cpg")
head(df_goldn %>% arrange(p))
```


***


## Quality Control Plots

For all cohorts, we want to plot:

* QQ plots side by side
* Volcano plots side by side
* Manhattan plots side by side
* Boxplots of effect sizes
* Boxplots of SEs

***

### QQ plots 


***

LLS

```{r}
draw_qq("lls")
```

***

KORA

```{r}
draw_qq("kora")
```

***

TwinsUK

```{r}
draw_qq("tuk")
```

***

LifeLines Deep

```{r}
draw_qq("lld")
```

***

SHIP

```{r}
draw_qq("ship")
```

***

GOLDN

```{r}
draw_qq("goldn")
```

***

Draw

```{r out.width="100%"}
ggarrange(plot_lls, 
          plot_kora, 
          plot_tuk,
          plot_lld,
          plot_ship,
          plot_goldn,
          ncol = 3, nrow = 2)
```

***

### Volcano Plots

***

Make scale

```{r}
min <- as.numeric(-max(c(
  abs(bacon_df$beta_lls), 
  abs(bacon_df$beta_kora),
  abs(bacon_df$beta_tuk),
  abs(bacon_df$beta_lld),
  abs(bacon_df$beta_ship),
  abs(bacon_df$beta_goldn)),
  na.rm=TRUE) - 0.005)

max <- as.numeric(max(c(
  abs(bacon_df$beta_lls), 
  abs(bacon_df$beta_kora),
  abs(bacon_df$beta_tuk),
  abs(bacon_df$beta_lld),
  abs(bacon_df$beta_ship),
  abs(bacon_df$beta_goldn)),
  na.rm=TRUE) + 0.005)

p_max <- as.numeric(-log10(min(c(
  bacon_df$p_lls,
  bacon_df$p_kora,
  bacon_df$p_tuk,
  bacon_df$p_lld,
  bacon_df$p_ship,
  bacon_df$p_goldn),
  na.rm=TRUE))) + 2
```

***

LLS

```{r}
draw_volcano("lls")
```

***

KORA

```{r}
draw_volcano("kora")
```

***

TwinsUK

```{r}
draw_volcano("tuk")
```

***

LifeLines Deep

```{r}
draw_volcano("lld")
```

***

SHIP

```{r}
draw_volcano("ship")
```

***

GOLDN

```{r}
draw_volcano("goldn")
```

***

Draw

```{r out.width="100%"}
ggarrange(
  plot_lls, 
  plot_kora, 
  plot_tuk,
  plot_lld,
  plot_ship,
  plot_goldn,
  ncol = 3, nrow = 2)
```

***

### Manhattan plots

***

LLS

```{r}
draw_man("lls")
```

***

KORA

```{r}
draw_man("kora")
```

***

TwinsUK

```{r}
draw_man("tuk")
```

***

LifeLines Deep

```{r}
draw_man("lld")
```

***

SHIP

```{r}
draw_man("ship")
```

***

GOLDN

```{r}
draw_man("goldn")
```

***

Draw

```{r out.width="100%"}
ggarrange(
  plot_lls, 
  plot_kora, 
  plot_tuk,
  plot_lld,
  plot_ship,
  plot_goldn,
  ncol = 3, nrow = 2)
```

***

### Boxplots

***

Make plot of effect size distributions in each cohort

```{r}
bacon_df <- bacon_df %>% 
  mutate(
    beta_LLS = as.numeric(beta_lls),
    beta_KORA = as.numeric(beta_kora),
    beta_TwinsUK = as.numeric(beta_tuk),
    beta_LLD = as.numeric(beta_lld),
    beta_SHIP = as.numeric(beta_ship),
    beta_GOLDN = as.numeric(beta_goldn)
  )

beta_df <- pivot_longer(
  data = bacon_df,
  cols = c("beta_LLS", "beta_KORA", "beta_TwinsUK", 
           "beta_LLD", "beta_SHIP", "beta_GOLDN"),
  names_to = "cohort",
  names_prefix = "beta_",
  values_to = "beta",
  values_drop_na = TRUE
)

beta_df %>% 
  ggplot(aes(x = cohort, y = beta, fill = cohort)) +
  scale_fill_manual(values = c('#e2455f', '#66cec8', '#eaa87d',
                               '#f9f5e0', '#794f7a', '#5FA052')) +
  geom_boxplot(alpha = 0.8) +
  ggtitle("Effect size distributions by cohort") +
  ylab("\u03b2") + xlab("Cohort") +
  theme( 
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021", fill=NA),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(size = 12, color = "#1B2021"),
    axis.text.x = element_text(size = 12, color = "#1B2021"),
    axis.title = element_text(size=11, color = "#1B2021"),
    plot.title = element_text(size = 16, hjust=0.5,
                              color = "#548687", face = "bold"))
```

***

Make plot of SE distribution in each cohort

```{r}
bacon_df <- bacon_df %>% 
  mutate(
    SE_LLS = as.numeric(SE_lls),
    SE_KORA = as.numeric(SE_kora),
    SE_TwinsUK = as.numeric(SE_tuk),
    SE_LLD = as.numeric(SE_lld),
    SE_SHIP = as.numeric(SE_ship),
    SE_GOLDN = as.numeric(SE_goldn)
  )

SE_df <- pivot_longer(
  data = bacon_df,
  cols = c("SE_LLS", "SE_KORA", "SE_TwinsUK", 
           "SE_LLD", "SE_SHIP", "SE_GOLDN"),
  names_to = "cohort",
  names_prefix = "SE_",
  values_to = "SE",
  values_drop_na = TRUE
)

SE_df %>% 
  ggplot(aes(x = cohort, y = SE, fill = cohort)) +
  scale_fill_manual(values = c('#e2455f', '#66cec8', '#eaa87d',
                               '#f9f5e0', '#794f7a', '#5FA052')) +
  geom_boxplot(alpha = 0.6) +
  ggtitle("SE distributions by cohort") +
  ylab("SE") + xlab("Cohort") +
  theme( 
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021", fill=NA),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(size = 12, color = "#1B2021"),
    axis.text.x = element_text(size = 12, color = "#1B2021"),
    axis.title = element_text(size=11, color = "#1B2021"),
    plot.title = element_text(size = 16, hjust=0.5,
                              color = "#548687", face = "bold"))
```


***

# Adiponectin EWAS - BMI Sensitivity Analysis

***

### Load data

This contains data on:

* `cpg` - probe ID
* `beta_` - effect size reported by each cohort
* `SE_` - standard error reported by each cohort
* `p_`- unadjusted p-value reported by each cohort
* `N_` - sample size of each cohort

```{r}
load('../Adi_Data/Processing/ALL-3-BACON_ln-adi_ext2.Rdata')
bacon_df[1:5, 1:5]
```

***

### Save Results

Create function

```{r}
save_results <- function(cohort){
  assign(paste0("df_", cohort),
         bacon_df %>% 
           dplyr::select(
             cpg,
             beta = paste0("beta_", cohort),
             se = paste0("SE_", cohort),
             p = paste0("p_", cohort),
             N = paste0("N_", cohort)
           ) %>% 
           mutate(
             beta = as.numeric(beta),
             se = as.numeric(se), 
             p = as.numeric(p),
             N = as.numeric(N) 
           ) %>% 
           filter(N>50),
         envir=.GlobalEnv)
}
```

***

LLS

```{r}
save_results("lls")
write_tsv(df_lls, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext2-LLS.tsv")
```

***

KORA

```{r}
save_results("kora")
write_tsv(df_kora, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext2-KORA.tsv")
```

***

TwinsUK

```{r}
save_results("tuk")
write_tsv(df_tuk, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext2-TUK.tsv")
```

***

LifeLines Deep

```{r}
save_results("lld")
write_tsv(df_lld, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext2-LLD.tsv")
```

***

SHIP

```{r}
save_results("ship")
write_tsv(df_ship, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext2-SHIP.tsv")
```

***

GOLDN

```{r}
save_results("goldn")
write_tsv(df_goldn, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext2-GOLDN.tsv")
```


### Top Hits

***

```{r}
df_lls <- left_join(df_lls, anno, by = "cpg")
head(df_lls %>% arrange(p))
```

***

KORA

```{r}
df_kora <- left_join(df_kora, anno, by = "cpg")
head(df_kora %>% arrange(p))
```

***

TwinsUK

```{r}
df_tuk <- left_join(df_tuk, anno, by = "cpg")
head(df_tuk %>% arrange(p))
```

***

LifeLines Deep

```{r}
df_lld <- left_join(df_lld, anno, by = "cpg")
head(df_lld %>% arrange(p))
```

***

SHIP

```{r}
df_ship <- left_join(df_ship, anno, by = "cpg")
head(df_ship %>% arrange(p))
```

***

GOLDN

```{r}
df_goldn <- left_join(df_goldn, anno, by="cpg")
head(df_goldn %>% arrange(p))
```


***


## Quality Control Plots

For all cohorts, we want to plot:

* QQ plots side by side
* Volcano plots side by side
* Manhattan plots side by side
* Boxplots of effect sizes
* Boxplots of SEs

***

### QQ plots 


***

LLS

```{r}
draw_qq("lls")
```

***

KORA

```{r}
draw_qq("kora")
```

***

TwinsUK

```{r}
draw_qq("tuk")
```

***

LifeLines Deep

```{r}
draw_qq("lld")
```

***

SHIP

```{r}
draw_qq("ship")
```

***

GOLDN

```{r}
draw_qq("goldn")
```

***

Draw

```{r out.width="100%"}
ggarrange(plot_lls, 
          plot_kora, 
          plot_tuk,
          plot_lld,
          plot_ship,
          plot_goldn,
          ncol = 3, nrow = 2)
```

***

### Volcano Plots

***

Make scale

```{r}
min <- as.numeric(-max(c(
  abs(bacon_df$beta_lls), 
  abs(bacon_df$beta_kora),
  abs(bacon_df$beta_tuk),
  abs(bacon_df$beta_lld),
  abs(bacon_df$beta_ship),
  abs(bacon_df$beta_goldn)),
  na.rm=TRUE) - 0.005)

max <- as.numeric(max(c(
  abs(bacon_df$beta_lls), 
  abs(bacon_df$beta_kora),
  abs(bacon_df$beta_tuk),
  abs(bacon_df$beta_lld),
  abs(bacon_df$beta_ship),
  abs(bacon_df$beta_goldn)),
  na.rm=TRUE) + 0.005)

p_max <- as.numeric(-log10(min(c(
  bacon_df$p_lls,
  bacon_df$p_kora,
  bacon_df$p_tuk,
  bacon_df$p_lld,
  bacon_df$p_ship,
  bacon_df$p_goldn),
  na.rm=TRUE))) + 2
```

***

LLS

```{r}
draw_volcano("lls")
```

***

KORA

```{r}
draw_volcano("kora")
```

***

TwinsUK

```{r}
draw_volcano("tuk")
```

***

LifeLines Deep

```{r}
draw_volcano("lld")
```

***

SHIP

```{r}
draw_volcano("ship")
```

***

GOLDN

```{r}
draw_volcano("goldn")
```

***

Draw

```{r out.width="100%"}
ggarrange(
  plot_lls, 
  plot_kora, 
  plot_tuk,
  plot_lld,
  plot_ship,
  plot_goldn,
  ncol = 3, nrow = 2)
```

***

### Manhattan plots

***

LLS

```{r}
draw_man("lls")
```

***

KORA

```{r}
draw_man("kora")
```

***

TwinsUK

```{r}
draw_man("tuk")
```

***

LifeLines Deep

```{r}
draw_man("lld")
```

***

SHIP

```{r}
draw_man("ship")
```

***

GOLDN

```{r}
draw_man("goldn")
```

***

Draw

```{r out.width="100%"}
ggarrange(
  plot_lls, 
  plot_kora, 
  plot_tuk,
  plot_lld,
  plot_ship,
  plot_goldn,
  ncol = 3, nrow = 2)
```

***

### Boxplots

***

Make plot of effect size distributions in each cohort

```{r}
bacon_df <- bacon_df %>% 
  mutate(
    beta_LLS = as.numeric(beta_lls),
    beta_KORA = as.numeric(beta_kora),
    beta_TwinsUK = as.numeric(beta_tuk),
    beta_LLD = as.numeric(beta_lld),
    beta_SHIP = as.numeric(beta_ship),
    beta_GOLDN = as.numeric(beta_goldn)
  )

beta_df <- pivot_longer(
  data = bacon_df,
  cols = c("beta_LLS", "beta_KORA", "beta_TwinsUK", 
           "beta_LLD", "beta_SHIP", "beta_GOLDN"),
  names_to = "cohort",
  names_prefix = "beta_",
  values_to = "beta",
  values_drop_na = TRUE
)

beta_df %>% 
  ggplot(aes(x = cohort, y = beta, fill = cohort)) +
  scale_fill_manual(values = c('#e2455f', '#66cec8', '#eaa87d',
                               '#f9f5e0', '#794f7a', '#5FA052')) +
  geom_boxplot(alpha = 0.8) +
  ggtitle("Effect size distributions by cohort") +
  ylab("\u03b2") + xlab("Cohort") +
  theme( 
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021", fill=NA),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(size = 12, color = "#1B2021"),
    axis.text.x = element_text(size = 12, color = "#1B2021"),
    axis.title = element_text(size=11, color = "#1B2021"),
    plot.title = element_text(size = 16, hjust=0.5,
                              color = "#548687", face = "bold"))
```

***

Make plot of SE distribution in each cohort

```{r}
bacon_df <- bacon_df %>% 
  mutate(
    SE_LLS = as.numeric(SE_lls),
    SE_KORA = as.numeric(SE_kora),
    SE_TwinsUK = as.numeric(SE_tuk),
    SE_LLD = as.numeric(SE_lld),
    SE_SHIP = as.numeric(SE_ship),
    SE_GOLDN = as.numeric(SE_goldn)
  )

SE_df <- pivot_longer(
  data = bacon_df,
  cols = c("SE_LLS", "SE_KORA", "SE_TwinsUK", 
           "SE_LLD", "SE_SHIP", "SE_GOLDN"),
  names_to = "cohort",
  names_prefix = "SE_",
  values_to = "SE",
  values_drop_na = TRUE
)

SE_df %>% 
  ggplot(aes(x = cohort, y = SE, fill = cohort)) +
  scale_fill_manual(values = c('#e2455f', '#66cec8', '#eaa87d',
                               '#f9f5e0', '#794f7a', '#5FA052')) +
  geom_boxplot(alpha = 0.6) +
  ggtitle("SE distributions by cohort") +
  ylab("SE") + xlab("Cohort") +
  theme( 
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021", fill=NA),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(size = 12, color = "#1B2021"),
    axis.text.x = element_text(size = 12, color = "#1B2021"),
    axis.title = element_text(size=11, color = "#1B2021"),
    plot.title = element_text(size = 16, hjust=0.5,
                              color = "#548687", face = "bold"))
```

***



# Adiponectin EWAS - Epidish Sensitivity Analysis

***

### Load data

This contains data on:

* `cpg` - probe ID
* `beta_` - effect size reported by each cohort
* `SE_` - standard error reported by each cohort
* `p_`- unadjusted p-value reported by each cohort
* `N_` - sample size of each cohort

```{r}
load('../Adi_Data/Processing/ALL-3-BACON_ln-adi_ext3.Rdata')
bacon_df[1:5, 1:5]
```

***

### Save Results

Create function

```{r}
save_results <- function(cohort){
  assign(paste0("df_", cohort),
         bacon_df %>% 
           dplyr::select(
             cpg,
             beta = paste0("beta_", cohort),
             se = paste0("SE_", cohort),
             p = paste0("p_", cohort),
             N = paste0("N_", cohort)
           ) %>% 
           mutate(
             beta = as.numeric(beta),
             se = as.numeric(se), 
             p = as.numeric(p),
             N = as.numeric(N) 
           ) %>% 
           filter(N>50),
         envir=.GlobalEnv)
}
```

***

LLS

```{r}
save_results("lls")
write_tsv(df_lls, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext3-LLS.tsv")
```

***

KORA

```{r}
save_results("kora")
write_tsv(df_kora, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext3-KORA.tsv")
```

***

LifeLines Deep

```{r}
save_results("lld")
write_tsv(df_lld, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext3-LLD.tsv")
```

***

SHIP

```{r}
save_results("ship")
write_tsv(df_ship, file="../Adi_Data/Processing/Meta_Input/OUT_adi_ext3-SHIP.tsv")
```

### Top Hits

***

```{r}
df_lls <- left_join(df_lls, anno, by = "cpg")
head(df_lls %>% arrange(p))
```

***

KORA

```{r}
df_kora <- left_join(df_kora, anno, by = "cpg")
head(df_kora %>% arrange(p))
```

***

LifeLines Deep

```{r}
df_lld <- left_join(df_lld, anno, by = "cpg")
head(df_lld %>% arrange(p))
```

***

SHIP

```{r}
df_ship <- left_join(df_ship, anno, by = "cpg")
head(df_ship %>% arrange(p))
```

***


## Quality Control Plots

For all cohorts, we want to plot:

* QQ plots side by side
* Volcano plots side by side
* Manhattan plots side by side
* Boxplots of effect sizes
* Boxplots of SEs

***

### QQ plots 


***

LLS

```{r}
draw_qq("lls")
```

***

KORA

```{r}
draw_qq("kora")
```

***

LifeLines Deep

```{r}
draw_qq("lld")
```

***

SHIP

```{r}
draw_qq("ship")
```

***

Draw

```{r out.width="100%"}
ggarrange(plot_lls, 
          plot_kora, 
          plot_lld,
          plot_ship,
          ncol = 2, nrow = 2)
```

***

### Volcano Plots

***

Make scale

```{r}
min <- as.numeric(-max(c(
  abs(bacon_df$beta_lls), 
  abs(bacon_df$beta_kora),
  abs(bacon_df$beta_lld),
  abs(bacon_df$beta_ship)),
  na.rm=TRUE) - 0.005)

max <- as.numeric(max(c(
  abs(bacon_df$beta_lls), 
  abs(bacon_df$beta_kora),
  abs(bacon_df$beta_lld),
  abs(bacon_df$beta_ship)),
  na.rm=TRUE) + 0.005)

p_max <- as.numeric(-log10(min(c(
  bacon_df$p_lls,
  bacon_df$p_kora,
  bacon_df$p_lld,
  bacon_df$p_ship),
  na.rm=TRUE))) + 2
```

***

LLS

```{r}
draw_volcano("lls")
```

***

KORA

```{r}
draw_volcano("kora")
```

***

LifeLines Deep

```{r}
draw_volcano("lld")
```

***

SHIP

```{r}
draw_volcano("ship")
```

***

Draw

```{r out.width="100%"}
ggarrange(
  plot_lls, 
  plot_kora, 
  plot_lld,
  plot_ship,
  ncol = 2, nrow = 2)
```

***

### Manhattan plots

***

LLS

```{r}
draw_man("lls")
```

***

KORA

```{r}
draw_man("kora")
```

***

LifeLines Deep

```{r}
draw_man("lld")
```

***

SHIP

```{r}
draw_man("ship")
```

***

Draw

```{r out.width="100%"}
ggarrange(
  plot_lls, 
  plot_kora, 
  plot_lld,
  plot_ship,
  ncol = 2, nrow = 2)
```

***

### Boxplots

***

Make plot of effect size distributions in each cohort

```{r}
bacon_df <- bacon_df %>% 
  mutate(
    beta_LLS = as.numeric(beta_lls),
    beta_KORA = as.numeric(beta_kora),
    beta_LLD = as.numeric(beta_lld),
    beta_SHIP = as.numeric(beta_ship)
  )

beta_df <- pivot_longer(
  data = bacon_df,
  cols = c("beta_LLS", "beta_KORA", 
           "beta_LLD", "beta_SHIP"),
  names_to = "cohort",
  names_prefix = "beta_",
  values_to = "beta",
  values_drop_na = TRUE
)

beta_df %>% 
  ggplot(aes(x = cohort, y = beta, fill = cohort)) +
  scale_fill_manual(values = c('#e2455f', '#66cec8', '#eaa87d',
                               '#f9f5e0')) +
  geom_boxplot(alpha = 0.8) +
  ggtitle("Effect size distributions by cohort") +
  ylab("\u03b2") + xlab("Cohort") +
  theme( 
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021", fill=NA),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(size = 12, color = "#1B2021"),
    axis.text.x = element_text(size = 12, color = "#1B2021"),
    axis.title = element_text(size=11, color = "#1B2021"),
    plot.title = element_text(size = 16, hjust=0.5,
                              color = "#548687", face = "bold"))
```

***

Make plot of SE distribution in each cohort

```{r}
bacon_df <- bacon_df %>% 
  mutate(
    SE_LLS = as.numeric(SE_lls),
    SE_KORA = as.numeric(SE_kora),
    SE_LLD = as.numeric(SE_lld),
    SE_SHIP = as.numeric(SE_ship)
  )

SE_df <- pivot_longer(
  data = bacon_df,
  cols = c("SE_LLS", "SE_KORA",  
           "SE_LLD", "SE_SHIP"),
  names_to = "cohort",
  names_prefix = "SE_",
  values_to = "SE",
  values_drop_na = TRUE
)

SE_df %>% 
  ggplot(aes(x = cohort, y = SE, fill = cohort)) +
  scale_fill_manual(values = c('#e2455f', '#66cec8', '#eaa87d',
                               '#f9f5e0')) +
  geom_boxplot(alpha = 0.6) +
  ggtitle("SE distributions by cohort") +
  ylab("SE") + xlab("Cohort") +
  theme( 
    panel.background = element_rect(
      fill="white"),
    panel.border = element_rect(
      color="#1B2021", fill=NA),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(size = 12, color = "#1B2021"),
    axis.text.x = element_text(size = 12, color = "#1B2021"),
    axis.title = element_text(size=11, color = "#1B2021"),
    plot.title = element_text(size = 16, hjust=0.5,
                              color = "#548687", face = "bold"))
```