---
title: "Adipokine EWAS - Script 8: Save Full Results"
author: "Lucy Sinke"
output: html_document
---

```{r echo=FALSE}
rm(list=ls())
```

***

# Setup

Load packages

```{r warning=FALSE, message=FALSE}
library(tidyverse)
```

***

# Meta-Analysis Results

Create function

```{r}
load_meta <- function(folder){
  results <- read_tsv(
    file=paste0("../Adi_Data/Processing/Meta_Output/",
                folder,
                "/METAANALYSIS1.TBL"))
  
  results <- results %>% 
    mutate(
      pFDR = p.adjust(
        `P-value`,
        method="fdr"),
      pBonf = p.adjust(
        `P-value`, 
        method="bonf")) %>% 
    dplyr::select(
      MarkerName, 
      Effect,
      StdErr,
      `P-value`,
      pFDR,
      pBonf,
      Direction,
      HetISq,
      HetPVal)
  
  return(results)
}
```

Base model

```{r}
results <- load_meta(
  folder="Adi_Base")

colnames(results) <- c(
  "cpg", "base_meta_beta", "base_meta_se",
  "base_meta_p", "base_meta_padj_fdr", "base_meta_padj_bonf",
  "base_het_dir", "base_het_i2", "base_het_p"
)

adi_results <- results
```

Smoking sensitivity analysis

```{r}
results <- load_meta(
  folder="Adi_Smoke")

colnames(results) <- c(
  "cpg", "ext1_meta_beta", "ext1_meta_se",
  "ext1_meta_p", "ext1_meta_padj_fdr", "ext1_meta_padj_bonf",
  "ext1_het_dir", "ext1_het_i2", "ext1_het_p"
)

adi_results <- full_join(adi_results, results)
```

BMI sensitivity analysis

```{r}
results <- load_meta(
  folder="Adi_BMI")

colnames(results) <- c(
  "cpg", "ext2_meta_beta", "ext2_meta_se",
  "ext2_meta_p", "ext2_meta_padj_fdr", "ext2_meta_padj_bonf",
  "ext2_het_dir", "ext2_het_i2", "ext2_het_p"
)

adi_results <- full_join(adi_results, results)
```

Epidish sensitivity analysis

```{r}
results <- load_meta(
  folder="Adi_Epidish")

colnames(results) <- c(
  "cpg", "ext3_meta_beta", "ext3_meta_se",
  "ext3_meta_p", "ext3_meta_padj_fdr", "ext3_meta_padj_bonf",
  "ext3_het_dir", "ext3_het_i2", "ext3_het_p"
)

adi_results <- full_join(adi_results, results)
```


Inspect output

```{r}
str(adi_results)
```

***

# Cohort Specific Results

Create function

```{r}
read_cohort <- function(cohort, model){
  results <- read_tsv(
    file=paste0("../Adi_Data/Processing/Meta_Input/OUT_adi_",
                model,
                "-",
                toupper(cohort),
                ".tsv"))
  
  results <- results %>% 
    mutate(
      pFDR = p.adjust(p, method="bonf"),
      pBonf = p.adjust(p, method="fdr")
    ) %>% 
    dplyr::select(
      cpg,
      N,
      beta,
      se,
      p,
      pFDR,
      pBonf)
  
  return(results)
}
```

***

### LLS

Base model

```{r}
results <- read_cohort("lls", "base")

colnames(results) <- c(
  "cpg", "base_N_LLS", "base_beta_LLS", "base_SE_LLS",
  "base_p_LLS", "base_padj_fdr_LLS", "base_padj_bonf_LLS")

adi_results <- left_join(adi_results, results)
```

Smoking sensitivity analysis

```{r}
results <- read_cohort("lls", "ext1")

colnames(results) <- c(
  "cpg", "ext1_N_LLS", "ext1_beta_LLS", "ext1_SE_LLS",
  "ext1_p_LLS", "ext1_padj_fdr_LLS", "ext1_padj_bonf_LLS")

adi_results <- left_join(adi_results, results)
```

BMI sensitivity analysis

```{r}
results <- read_cohort("lls", "ext2")

colnames(results) <- c(
  "cpg", "ext2_N_LLS", "ext2_beta_LLS", "ext2_SE_LLS",
  "ext2_p_LLS", "ext2_padj_fdr_LLS", "ext2_padj_bonf_LLS")

adi_results <- left_join(adi_results, results)
```

BMI sensitivity analysis

```{r}
results <- read_cohort("lls", "ext3")

colnames(results) <- c(
  "cpg", "ext3_N_LLS", "ext3_beta_LLS", "ext3_SE_LLS",
  "ext3_p_LLS", "ext3_padj_fdr_LLS", "ext3_padj_bonf_LLS")

adi_results <- left_join(adi_results, results)
```


***

### KORA

Base model

```{r}
results <- read_cohort("kora", "base")

colnames(results) <- c(
  "cpg", "base_N_KORA", "base_beta_KORA", "base_SE_KORA",
  "base_p_KORA", "base_padj_fdr_KORA", "base_padj_bonf_KORA")

adi_results <- left_join(adi_results, results)
```

Smoking sensitivity analysis

```{r}
results <- read_cohort("kora", "ext1")

colnames(results) <- c(
  "cpg", "ext1_N_KORA", "ext1_beta_KORA", "ext1_SE_KORA",
  "ext1_p_KORA", "ext1_padj_fdr_KORA", "ext1_padj_bonf_KORA")

adi_results <- left_join(adi_results, results)
```

BMI sensitivity analysis

```{r}
results <- read_cohort("kora", "ext2")

colnames(results) <- c(
  "cpg", "ext2_N_KORA", "ext2_beta_KORA", "ext2_SE_KORA",
  "ext2_p_KORA", "ext2_padj_fdr_KORA", "ext2_padj_bonf_KORA")

adi_results <- left_join(adi_results, results)
```

BMI sensitivity analysis

```{r}
results <- read_cohort("kora", "ext3")

colnames(results) <- c(
  "cpg", "ext3_N_KORA", "ext3_beta_KORA", "ext3_SE_KORA",
  "ext3_p_KORA", "ext3_padj_fdr_KORA", "ext3_padj_bonf_KORA")

adi_results <- left_join(adi_results, results)
```


***

### LL Deep

Base model

```{r}
results <- read_cohort("lld", "base")

colnames(results) <- c(
  "cpg", "base_N_LLD", "base_beta_LLD", "base_SE_LLD",
  "base_p_LLD", "base_padj_fdr_LLD", "base_padj_bonf_LLD")

adi_results <- left_join(adi_results, results)
```

Smoking sensitivity analysis

```{r}
results <- read_cohort("lld", "ext1")

colnames(results) <- c(
  "cpg", "ext1_N_LLD", "ext1_beta_LLD", "ext1_SE_LLD",
  "ext1_p_LLD", "ext1_padj_fdr_LLD", "ext1_padj_bonf_LLD")

adi_results <- left_join(adi_results, results)
```

BMI sensitivity analysis

```{r}
results <- read_cohort("lld", "ext2")

colnames(results) <- c(
  "cpg", "ext2_N_LLD", "ext2_beta_LLD", "ext2_SE_LLD",
  "ext2_p_LLD", "ext2_padj_fdr_LLD", "ext2_padj_bonf_LLD")

adi_results <- left_join(adi_results, results)
```

Epidish sensitivity analysis

```{r}
results <- read_cohort("lld", "ext3")

colnames(results) <- c(
  "cpg", "ext3_N_LLD", "ext3_beta_LLD", "ext3_SE_LLD",
  "ext3_p_LLD", "ext3_padj_fdr_LLD", "ext3_padj_bonf_LLD")

adi_results <- left_join(adi_results, results)
```

***

### SHIP

Base model

```{r}
results <- read_cohort("ship", "base")

colnames(results) <- c(
  "cpg", "base_N_SHIP", "base_beta_SHIP", "base_SE_SHIP",
  "base_p_SHIP", "base_padj_fdr_SHIP", "base_padj_bonf_SHIP")

adi_results <- left_join(adi_results, results)
```

Smoking sensitivity analysis

```{r}
results <- read_cohort("ship", "ext1")

colnames(results) <- c(
  "cpg", "ext1_N_SHIP", "ext1_beta_SHIP", "ext1_SE_SHIP",
  "ext1_p_SHIP", "ext1_padj_fdr_SHIP", "ext1_padj_bonf_SHIP")

adi_results <- left_join(adi_results, results)
```

BMI sensitivity analysis

```{r}
results <- read_cohort("ship", "ext2")

colnames(results) <- c(
  "cpg", "ext2_N_SHIP", "ext2_beta_SHIP", "ext2_SE_SHIP",
  "ext2_p_SHIP", "ext2_padj_fdr_SHIP", "ext2_padj_bonf_SHIP")

adi_results <- left_join(adi_results, results)
```

Epidish sensitivity analysis

```{r}
results <- read_cohort("ship", "ext3")

colnames(results) <- c(
  "cpg", "ext3_N_SHIP", "ext3_beta_SHIP", "ext3_SE_SHIP",
  "ext3_p_SHIP", "ext3_padj_fdr_SHIP", "ext3_padj_bonf_SHIP")

adi_results <- left_join(adi_results, results)
```


***

### TwinsUK

Base model

```{r}
results <- read_cohort("tuk", "base")

colnames(results) <- c(
  "cpg", "base_N_TUK", "base_beta_TUK", "base_SE_TUK",
  "base_p_TUK", "base_padj_fdr_TUK", "base_padj_bonf_TUK")

adi_results <- left_join(adi_results, results)
```

Smoking sensitivity analysis

```{r}
results <- read_cohort("tuk", "ext1")

colnames(results) <- c(
  "cpg", "ext1_N_TUK", "ext1_beta_TUK", "ext1_SE_TUK",
  "ext1_p_TUK", "ext1_padj_fdr_TUK", "ext1_padj_bonf_TUK")

adi_results <- left_join(adi_results, results)
```

BMI sensitivity analysis

```{r}
results <- read_cohort("tuk", "ext2")

colnames(results) <- c(
  "cpg", "ext2_N_TUK", "ext2_beta_TUK", "ext2_SE_TUK",
  "ext2_p_TUK", "ext2_padj_fdr_TUK", "ext2_padj_bonf_TUK")

adi_results <- left_join(adi_results, results)
```


***

### GOLDN

Base model

```{r}
results <- read_cohort("goldn", "base")

colnames(results) <- c(
  "cpg", "base_N_GOLDN", "base_beta_GOLDN", "base_SE_GOLDN",
  "base_p_GOLDN", "base_padj_fdr_GOLDN", "base_padj_bonf_GOLDN")

adi_results <- left_join(adi_results, results)
```

Smoking sensitivity analysis

```{r}
results <- read_cohort("goldn", "ext1")

colnames(results) <- c(
  "cpg", "ext1_N_GOLDN", "ext1_beta_GOLDN", "ext1_SE_GOLDN",
  "ext1_p_GOLDN", "ext1_padj_fdr_GOLDN", "ext1_padj_bonf_GOLDN")

adi_results <- left_join(adi_results, results)
```

BMI sensitivity analysis

```{r}
results <- read_cohort("goldn", "ext2")

colnames(results) <- c(
  "cpg", "ext2_N_GOLDN", "ext2_beta_GOLDN", "ext2_SE_GOLDN",
  "ext2_p_GOLDN", "ext2_padj_fdr_GOLDN", "ext2_padj_bonf_GOLDN")

adi_results <- left_join(adi_results, results)
```

***

# Methylation Statistics

Load data

```{r}
load("../Adi_Data/Cohort_Results/ALL_ln-adi_base.Rdata")

limma_base <- limma_base %>% 
  dplyr::select(
    cpg, 
    starts_with("cpg_mean"),
    starts_with("cpg_sd")
  )

adi_results <- left_join(
  adi_results, limma_base,
  by="cpg"
)
```

***

# Annotation

***

### Variables of Interest

`manifest_hg19` (fetched on 4/4/2023 from https://zwdzwd.github.io/InfiniumAnnotation)

* `probeID` as `cpg` - CpG ID
* `CpG_chrm` as `cpg_chr_hg19` - chromosome (hg19)
* `CpG_beg` as `cpg_start_hg19` - CpG start position (hg19)
* `CpG_end` as `cpg_end_hg19` - CpG end position (hg19)
* `probe_strand` as `cpg_strand` - strand
* `gene_HGNC`

```{r}
manifest_hg19 <- read_tsv(
  "../Shared_Data/Manifests/HM450.hg19.manifest.tsv.gz")

anno <- manifest_hg19 %>% 
  dplyr::select(
    cpg = probeID,
    cpg_chr_hg19 = CpG_chrm,
    cpg_start_hg19 = CpG_beg,
    cpg_end_hg19 = CpG_end,
    cpg_strand = probe_strand,
    gene_HGNC
  ) %>% 
  mutate(
    cpg_chr_hg19 = substr(cpg_chr_hg19,4,5)
  )

anno <- anno %>% 
  dplyr::filter(cpg %in% adi_results$cpg)
```

***

`manifest_hg38` (fetched on 4/4/2023 from https://zwdzwd.github.io/InfiniumAnnotation)

* `Probe_ID` as `cpg` - CpG ID
* `CpG_chrm` as `cpg_chr_hg38` - chromosome (hg38)
* `CpG_beg` as `cpg_start_hg38` - CpG start position (hg38)
* `CpG_end` as `cpg_end_hg38` - CpG end position (hg38)

```{r}
manifest_hg38 <- read_tsv(
  "../Shared_Data/Manifests/HM450.hg38.manifest.tsv.gz"
)

manifest_hg38 <- manifest_hg38 %>% 
  dplyr::select(
    cpg = Probe_ID,
    cpg_chr_hg38 = CpG_chrm,
    cpg_start_hg38 = CpG_beg,
    cpg_end_hg38 = CpG_end
  ) %>% 
  mutate(
    cpg_chr_hg38 = substr(cpg_chr_hg38,4,5)
  )

anno <- left_join(
  anno, manifest_hg38,
  by="cpg"
)
```

***

`manifest_roadmap` (fetched on 4/4/2023 from https://zwdzwd.github.io/InfiniumAnnotation)

* `probeID` as `cpg` - CpG ID
* `E062` - ROADMAP PBMC chromatin state annotation

```{r}
manifest_roadmap <- read_tsv(
  "../Shared_Data/Manifests/HM450.hg19.REMC.chromHMM.tsv.gz"
)

manifest_roadmap <- manifest_roadmap %>% 
  dplyr::select(
    cpg = probeID,
    E062
  )

anno <- left_join(
  anno, manifest_roadmap,
  by="cpg"
)

adi_results <- left_join(
  adi_results, anno
)
```

***

# Heterogeneity

Look at heterogenous CpGs

```{r}
het <- adi_results %>% filter(base_het_i2 > 90)
het_sig <- het %>% filter(base_meta_padj_fdr <= 0.05)
print(paste0('There are ', nrow(het), ' heterogenous CpGs that will be removed, and ', nrow(het_sig), ' of these were significant...'))
```

***

# Saving

Full results

```{r}
save(adi_results, 
     file="../Adi_Data/Results/Adi_EWAS_Results-Full.Rdata")
```

***

